Non-word repetition in Yélî Dnye

### 2020-05-20 manuscript.Rmd
advanced a bit

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

@@

- add ana based on phonemes & expand on levenshtein
- add ana based on number of characters
- add summary of types of error: %semantic, %deletion (which sounds?) %subst (which are the most common sounds) 
- maybe add to the last ana the prevalence (ie out of all the gh which proportion get transformed)
- add analysis on the basis of demo chars
- discussion

@@

#remove children who do not have both first & subsequent attempts
pwcm_att[pwcm_att$ID %in% names(table(pwcm_att$ID))[table(pwcm_att$ID)==2],]->pwcm_att

@@

We first checked whether accuracy varies between first and subsequent presentations of an item by averaging word-level accuracy at the participant level separately for first attempts and subsequent repetitions. As shown in Figure 1, participants' mean word-level accuracy became more heterogeneous in subsequent repetitions. Surprisingly, subsequent repetitions (M = `r round(mean(pwcm_att$Word[pwcm_att$Attempt!="first"]))`, SD = `r round(sd(pwcm_att$Word[pwcm_att$Attempt!="first"]))`)
were on average less accurate than first ones (M = `r round(mean(pwcm_att$Word[pwcm_att$Attempt=="first"]))`, SD = `r round(sd(pwcm_att$Word[pwcm_att$Attempt=="first"]))`), t(`r round(myt$parameter,2)`) = `r round(myt$statistic,2)`, p = `r round(myt$p.value,2)`. Given the uncertainty in whether previous work used only the first or all repetitions, and since behavior degraded and became more heterogeneous in subsequent repetitions, the rest of the analyses focus on only the first repetitions.

```{r exclude subsequent}
trnsc=trnsc[trnsc$attempt=="first",]
```

@@

Taking into account only the first attempts, the overall repetition accuracy measured in whole words is `r round(mean(trnsc$cor.online)*100)`% based on the online coding (i.e., whether the native experimenter asked for the target to be re-played), and `r round(mean(trnsc$correct)*100)`% when the offline coding is considered. On average, the phoneme-based normalized Levenshtein distance was `r round(mean(trnsc$nld)*100)`%, meaning that about 20% of phonemes were substituted, inserted, or deleted.

ADD GRAPH WITH PHONEME BASED IN Y AND WORD BASED IN X

@@

We then inspected whether accuracy varied as a function of word length.

```{r figlength,eval=F}
#not working
aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$age.exact,trnsc$tarlen),mean)->pwcm_att
colnames(pwcm_att)<-c("ID","Age","Length","Word")

#but this one is! What a mystery...
aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$age.exact,trnsc$type),mean)->pwcm_att
colnames(pwcm_att)<-c("ID","Age","Length","Word")

myx=as.numeric(as.factor(pwcm_att$ID))/100+as.numeric(as.factor(pwcm_att$Length))

plot(pwcm_att$Word~myx,pch=20,cex=2,lty=3,xlab="",xaxt="n",ylab="NWR scores (%)")
axis(1,at=1:length(levels(factor(pwcm_att$Length))),labels=levels(factor(pwcm_att$Length)))

@@

>>>>

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

@@

- add table r/d by stim length & age
- add table with stimuli
- ask Middy about consent, correct, accuracy
- double check demo: exact age missing: 78, 82, 83, 115, 119; rough age missing 78, there's a mixture of BGMF in sex, first born is missing, round age
- add summary of types of error: %semantic, %deletion (which sounds?) %subst (which are the most common sounds) 
- maybe add to the last ana the prevalence (ie out of all the gh which proportion get transformed)
- add analysis on the basis of demo chars
- discussion

@@

#remove children who do not have both first & subsequent attempts
# which is
excl=sum(table(pwcm_att$ID)!=2)

pwcm_att[pwcm_att$ID %in% names(table(pwcm_att$ID))[table(pwcm_att$ID)==2],]->pwcm_att

@@

We first checked whether accuracy varies between first and subsequent presentations of an item by averaging word-level accuracy at the participant level separately for first attempts and subsequent repetitions. We excluded `r excl` children who did not have data for one of these two types. As shown in Figure 1, participants' mean word-level accuracy became more heterogeneous in subsequent repetitions. Surprisingly, subsequent repetitions (M = `r round(mean(pwcm_att$Word[pwcm_att$Attempt!="first"]))`, SD = `r round(sd(pwcm_att$Word[pwcm_att$Attempt!="first"]))`)
were on average less accurate than first ones (M = `r round(mean(pwcm_att$Word[pwcm_att$Attempt=="first"]))`, SD = `r round(sd(pwcm_att$Word[pwcm_att$Attempt=="first"]))`), t(`r round(myt$parameter,2)`) = `r round(myt$statistic,2)`, p = `r round(myt$p.value,2)`. Given the uncertainty in whether previous work used only the first or all repetitions, and since behavior degraded and became more heterogeneous in subsequent repetitions, the rest of the analyses focus on only the first repetitions.

```{r exclude subsequent}
trnsc=trnsc[trnsc$attempt=="first",]

@@

```{r overall}

aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$age.exact),mean)->overall
aggregate(trnsc$pc*100,by=list(trnsc$id,trnsc$age.exact),mean)->pc
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$age.exact),mean)->nld
overall=cbind(overall,pc$x,nld$x)
colnames(overall)<-c("ID","Age","Word","Phoneme","NLD")

Taking into account only the first attempts, we averaged attempts by each of the `r dim(overall)[1]` children who had data for first attempts; their ages ranged from `r min(overall$Age)` to `r max(overall$Age)` (M = `r round(mean(overall$Age),1)`, SD = `r round(sd(overall$Age),1)`).

The overall NWR score was M = `r round(mean(overall$Word))`% (SD = `r round(sd(overall$Word))`%). Scores based on phonemes are even higher  M = `r round(mean(overall$Phoneme))`% (SD = `r round(sd(overall$Phoneme))`%). The phoneme-based normalized Levenshtein distance was  M = `r round(mean(overall$NLD))`% (SD = `r round(sd(overall$NLD))`%), meaning that about a fifth of phonemes were were substituted, inserted, or deleted. Notice that the normalized Levenshtein distance is the complement of phoneme-based scores.

@@

```{r bylength}
aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$age.exact,trnsc$type),mean)->by_length
aggregate(trnsc$pc*100,by=list(trnsc$id,trnsc$age.exact,trnsc$type),mean)->pc
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$age.exact,trnsc$type),mean)->nld
by_length=cbind(by_length,pc$x,nld$x)
colnames(by_length)<-c("ID","Age","Length","Word","Phoneme","NLD")

@@

We then inspected whether accuracy varied as a function of word length. Results are shown on table XX. This table shows that monosyllables accuracy was much lower than other lengths. This is likely due to the fact that the majority of monosyllables included were chosen because they had sounds that are rare in the world's languages, which may indicate that they are hard to produce or to perceive.

Setting monosyllables aside, we observe the typical pattern of decreased accuracy for longer items, although this is particularly salient for the whole word scoring. This is the most commonly reported type of score, but it is also the least forgiving. The pattern is less marked when other two scores are used, which are less sensitive to errors.

```{r tablength}
mytab=NULL
for(myvar in c("Word","Phoneme","NLD")){
  mytab=cbind(mytab,
              paste0(round(aggregate(by_length[,myvar],by=list(by_length$Length),mean)$x),
              " (",round(aggregate(by_length[,myvar],by=list(by_length$Length),sd)$x),")"))
}
rownames(mytab)<-levels(by_length$Length)
colnames(mytab)<-c("Word","Phoneme","NLD")
apa_table(mytab,caption="NWR scores measured in whole word accuracy, phoneme-based accuracy, and normalized Levenshtein Distance, separately for the four stimuli lengths.")
```

### 2020-05-20 manuscript.tex
advanced a bit

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[english,,man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Non-word repetition in Yélî Dnye},
            pdfkeywords={phonology, non-word repetition, development},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

Correspondence concerning this article should be addressed to Alejandrina Cristia, 29, rue d’Ulm, 75005 Paris, France. E-mail: alecristia@gmail.com}
\keywords{phonology, non-word repetition, development\newline\indent Word count: ~xxx words}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

@@

\begin{itemize}
\tightlist
\item
  fix yaml
\item
  add ana based on phonemes \& expand on levenshtein
\item
  add ana based on number of characters
\item
  add summary of types of error: \%semantic, \%deletion (which sounds?) \%subst (which are the most common sounds)
\item
  maybe add to the last ana the prevalence (ie out of all the gh which proportion get transformed)
\item
  add analysis on the basis of demo chars
\item
  discussion
\end{itemize}

@@

We first checked whether accuracy varies between first and subsequent presentations of an item by averaging word-level accuracy at the participant level separately for first attempts and subsequent repetitions. As shown in Figure 1, participants' mean word-level accuracy became more heterogeneous in subsequent repetitions. Surprisingly, subsequent repetitions (M = 42, SD = 28)
were on average less accurate than first ones (M = 64, SD = 15), t(36) = 5.32, p = 0. Given the uncertainty in whether previous work used only the first or all repetitions, and since behavior degraded and became more heterogeneous in subsequent repetitions, the rest of the analyses focus on only the first repetitions.

@@

Taking into account only the first attempts, the overall repetition accuracy measured in whole words is 72\% based on the online coding (i.e., whether the native experimenter asked for the target to be re-played), and 65\% when the offline coding is considered. On average, the phoneme-based normalized Levenshtein distance was 22\%, meaning that about 20\% of phonemes were substituted, inserted, or deleted.

ADD GRAPH WITH PHONEME BASED IN Y AND WORD BASED IN X

@@

We then inspected whether accuracy varied as a function of word length.

>>>>

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[english,,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Non-word repetition in Yélî Dnye},
            pdfkeywords={phonology, non-word repetition, development},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

Correspondence concerning this article should be addressed to Alejandrina Cristia, 29, rue d’Ulm, 75005 Paris, France. E-mail: alecristia@gmail.com}
\keywords{phonology, non-word repetition, development\newline\indent Word count: ~xxx words}
\usepackage{lineno}

@@

\begin{itemize}
\tightlist
\item
  add table r/d by stim length \& age
\item
  add table with stimuli
\item
  ask Middy about consent, correct, accuracy
\item
  double check demo: exact age missing: 78, 82, 83, 115, 119; rough age missing 78, there's a mixture of BGMF in sex, first born is missing, round age
\item
  add summary of types of error: \%semantic, \%deletion (which sounds?) \%subst (which are the most common sounds)
\item
  maybe add to the last ana the prevalence (ie out of all the gh which proportion get transformed)
\item
  add analysis on the basis of demo chars
\item
  discussion
\end{itemize}

@@

We first checked whether accuracy varies between first and subsequent presentations of an item by averaging word-level accuracy at the participant level separately for first attempts and subsequent repetitions. We excluded 5 children who did not have data for one of these two types. As shown in Figure 1, participants' mean word-level accuracy became more heterogeneous in subsequent repetitions. Surprisingly, subsequent repetitions (M = 42, SD = 28)
were on average less accurate than first ones (M = 64, SD = 15), t(36) = 5.32, p = 0. Given the uncertainty in whether previous work used only the first or all repetitions, and since behavior degraded and became more heterogeneous in subsequent repetitions, the rest of the analyses focus on only the first repetitions.

@@

Taking into account only the first attempts, we averaged attempts by each of the 37 children who had data for first attempts; their ages ranged from 3.92 to 12.86 (M = 7.10, SD = 1.90).

The overall NWR score was M = 64\% (SD = 15\%). Scores based on phonemes are even higher M = 78\% (SD = 9\%). The phoneme-based normalized Levenshtein distance was M = 22\% (SD = 9\%), meaning that about a fifth of phonemes were were substituted, inserted, or deleted. Notice that the normalized Levenshtein distance is the complement of phoneme-based scores.

@@

We then inspected whether accuracy varied as a function of word length. Results are shown on table XX. This table shows that monosyllables accuracy was much lower than other lengths. This is likely due to the fact that the majority of monosyllables included were chosen because they had sounds that are rare in the world's languages, which may indicate that they are hard to produce or to perceive.

Setting monosyllables aside, we observe the typical pattern of decreased accuracy for longer items, although this is particularly salient for the whole word scoring. This is the most commonly reported type of score, but it is also the least forgiving. The pattern is less marked when other two scores are used, which are less sensitive to errors.
